---
title: "Classfication Analysis"
author: "Mohammed Ali"
date: "January 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table) #reading table
library(dplyr)
library(class)
```

# k-Nearest Neighbors (kNN)

## Recognizing a road sign with kNN
After several trips with a human behind the wheel, it is time for the self-driving car to attempt the test course alone.

```{r load_data}
signs <- fread("data/knn_traffic_signs.csv", stringsAsFactors = FALSE,
               showProgress = FALSE)
train_signs <- signs[sample == "train"]
test_signs <- signs[sample == "test"]
glimpse(signs)
```

Create a vector of sign labels to use with kNN by extracting the column 
```{r sign_types}
# Create a vector of labels
sign_types <- train_signs$sign_type
```

Identify the next_sign using the `knn()` function:

* Set the `train` argument equal to the `signs` data frame without the first column.
* Set the `test` argument equal to the data frame `next_sign`.
* Use the vector of labels `cl` argument.

```{r first_knn_classifier}
# Classify the next sign observed
knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[1:1, -c("sign_type", "id", "sample")], cl = sign_types)
```


## Exploring the traffic sign dataset
To better understand how the `knn()` function was able to classify the stop sign, it may help to examine the training dataset it used.

Each previously observed street sign was divided into a 4x4 grid, and the red, green, and blue level for each of the 16 center.

```{r count_sign_types}
# Count the number of signs of each type
table(train_signs$sign_type)
```

Run the provided `aggregate()` command to see whether the average red level might vary by sign type.
```{r aggregate}
# Check r10's average red level by sign type
aggregate(r10 ~ sign_type, data = train_signs, mean)
```


## Classifying a collection of road signs
Now that the autonomous vehicle has successfully stopped on its own, your team feels confident allowing the car to continue the test course.

```{r predict_all}
# Use kNN to identify the test road signs
signs_pred <- knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[, -c("sign_type", "id", "sample")], cl = sign_types)


# Create a confusion matrix of the actual versus predicted values
signs_actual <- test_signs$sign_type
table(signs_pred, signs_actual)

# Compute the accuracy
mean(signs_pred == signs_actual)
```

## Testing other 'k' values

By default, the `knn()` function in the class package uses only the single nearest neighbor.

Setting a `k` parameter allows the algorithm to consider additional nearby neighbors. This enlarges the collection of neighbors which will vote on the predicted class.

Compare `k` values of _1_, _7_, and _15_ to examine the impact on traffic sign classification accuracy.

```{r different_ks}
# Compute the accuracy of the baseline model (default k = 1)
k_1 <- knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[, -c("sign_type", "id", "sample")], cl = sign_types)
mean(signs_actual == k_1)

# Modify the above to set k = 7
k_7 <- knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[, -c("sign_type", "id", "sample")], cl = sign_types, k = 7)
mean(signs_actual == k_7)

# Set k = 15 and compare to the above
k_15 <- knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[, -c("sign_type", "id", "sample")], cl = sign_types, k = 15)
mean(signs_actual == k_15)
```

## Seeing how the neighbors voted
When multiple nearest neighbors hold a vote, it can sometimes be useful to examine whether the voters were unanimous or widely separated.

For example, knowing more about the voters' confidence in the classification could allow an autonomous vehicle to use caution in the case there is any chance at all that a stop sign is ahead.

Build a kNN model with the `prob = TRUE` parameter to compute the vote proportions. Set `k = 7`.
```{r prob}
# Use the prob parameter to get the proportion of votes for the winning class
sign_pred <- knn(train = train_signs[, -c("sign_type", "id", "sample")], test = test_signs[, -c("sign_type", "id", "sample")], cl = sign_types, k = 7, prob = TRUE)

# Get the "prob" attribute from the predicted classes
sign_prob <- attr(sign_pred, "prob")

# Examine the first several predictions
head(sign_pred)

# Examine the proportion of votes for the winning class
head(sign_prob)
```
